What We Eval in the Shadows

### 2021-04-15 paper/main.tex
Usage of eval

\mypara{Variable lookup and values.}
This corresponds to minimized expressions \c{X}, \c{V}. We also include here lookup to slots in a list or data frame, \ie minimized expression \c{\$}.
It is the predominant usage of \eval in Base, through function \c{match.args}. In CRAN, it correlates with more exotic environments passed to \eval.
The dominant site in numbers of calls in CRAN, in the custom object system of \emph{ggplot2}, \c{ggproto}, also evaluates a variable \c{`\_inherit`} in the definition environment of the class. It leads to a large diversity of environments, many being package environment or package environments wrapped in \c{new}.

Package \emph{statnet.common} handles formulas, R object that associates an expression of the form \c{x ~ expr} and a carry-on environment where to evaluate the formula. It evaluates the left-hand side of the formula, which is always a variable, in \c{eval\_lhs.formula}, as \c{eval(object[[2L]],envir=environment(object))} where \c{object} is the formula and \c{environment} returns the formula associated to the object.

Slot lookup happens very often to evaluate some of the elements of a call list returned by \c{match.call}.
For instance in the \emph{pwrFDR} package:
\begin{lstlisting}
m <- match.call()
...
average.power <- m$average.power <- eval(m$average.power, sys.parent())
\end{lstlisting}

@@

@@

@@

A related pattern is not to change the call but rather to evaluate the arguments in the parent frame. This could also be achieved with \c{substitute}, but \c{match.call} does it in bulk.
In package \emph{MedDietCalc},  in function \c{computeMDS95}, \c{match.call} retrieves the argument expressions of the call, which are evaluated in a dataframe \c{data} which contains the numeric data. Each argument of function \c{computeMDS95} is also a column in dataframe \c{data}. The advantage of using \c{eval} here is to be able to express an argument as a combination of several columns in the dataframe.
\begin{lstlisting}
computeMDS95 <- function (data, Vegetables, Legumes, FruitAndNuts, Cereals,
	 Potatoes = NULL, Meat, Dairy, 	Alcohol, Fats = NULL, MUFA = NULL,
	  SFA = NULL, Sex, men = "male", women = "female", 	frequency = NULL,
	  output = "percent", rm.na = FALSE) {

	arguments <- as.list( match.call())
	Vegetables <- eval(arguments$Vegetables, data)
	Legumes <- eval(arguments$Legumes, data)
	FruitAndNuts <- eval(arguments$FruitAndNuts, data)
	Cereals <- eval(arguments$Cereals, data)
	Potatoes <- eval(arguments$Potatoes, data)
	...
\end{lstlisting}

@@

\mypara{Debloating.} \todo{Wouldn't reducing boilerplate be a better name?} \Eval is often used as a means to reduce boilerplate code;
simple and repetitive code can easily be replaced with judicious use of \eval.
For example, the \c{data.table} package uses \eval to calls the \c{options}
function with named arguments taken from a vector of strings. While the benefits
are limited in this example, it is an attractive tool for programmers.
\begin{lstlisting}
  opts = c("datatable.verbose"="FALSE", # ...many others
  for (i in names(opts))
    eval(parse(text=paste0("options(",i,"=",opts[i],")")))
\end{lstlisting}
This pattern is a special case of code generation, recognizable by the
fact that \eval is executed in a loop.

\mypara{Trivial.} When values are passed to \eval, they are returned
unchanged. They are an example of trivial uses of \eval. Another
trivial use is the empty expression, often found in JavaScript, but
rare in R. \todo{Given that we do not have any data I would get rid of it}
\todo{We have the normalized expressions V and the small study I did on values passed to eval. And I did not find any eval() (empty) in the dataset. }

% Creating macros?

>>>>

\subsection{Minimized expressions}

\mypara{Variable lookup and values.}
This corresponds to minimized expressions \c{X}, \c{V}. We also include here lookup to slots in a list or data frame, \ie minimized expression \c{\$}.
It is the predominant usage of \eval in Base, through function \c{match.args}. In CRAN, it correlates with more exotic environments passed to \eval. \todo{THis is with a V in practice.}
The dominant site in numbers of calls in CRAN, in the custom object system of \emph{ggplot2}, \c{ggproto}, also evaluates a variable \c{`\_inherit`} in the definition environment of the class. It leads to a large diversity of environments, many being package environments or package environments wrapped in \c{new}.
\todo{Explain why they use eval and not get... because of \c{::}}

Package \emph{statnet.common} handles formulas, R object that associates an expression of the form \c{x~\textasciitilde~expr} and a carry-on environment where to evaluate the formula. It evaluates the left-hand side of the formula, which is always a variable, in \c{eval\_lhs.formula}, as \c{eval(object[[2L]],envir=environment(object))} where \c{object} is the formula and \c{environment} returns the formula associated to the object.

\todo{Show an actual example with \$}
Slot lookup happens very often to evaluate some of the elements of a call list returned by \c{match.call}.
For instance in the \emph{pwrFDR} package:
\begin{lstlisting}
m <- match.call()
...
average.power <- m$average.power <- eval(m$average.power, sys.parent())
\end{lstlisting}

@@

\todo{Add testthat or withr::defer\_handler (fancy on.exit)}

@@

\subsection{High-level patterns.}

@@

A related pattern is not to change the call but rather to evaluate the arguments in the parent frame. This could also be achieved with \c{substitute}, but \c{match.call} does it in bulk.
In package \emph{MedDietCalc},  in function \c{computeMDS95}, \c{match.call} retrieves the argument expressions of the call, which are evaluated in a dataframe \c{data} which contains the numeric data. Each argument of function \c{computeMDS95} is also a column in dataframe \c{data}. The advantage of using \c{eval} here is to be able to express an argument as a combination of several columns in the dataframe.
\begin{lstlisting}
computeMDS95 <- function (data, Vegetables, ...) {
	arguments <- as.list( match.call())
	Vegetables <- eval(arguments$Vegetables, data) 
\end{lstlisting}

\todo{Put a simpler example.}

@@

 Code generation is often used as a means to reduce boilerplate code;
simple and repetitive code can easily be replaced with judicious use of \eval.
For example, the \c{data.table} package uses \eval to calls the \c{options}
function with named arguments taken from a vector of strings. While the benefits
are limited in this example, it is an attractive tool for programmers.
\begin{lstlisting}
  opts = c("datatable.verbose"="FALSE", # ...many others
  for (i in names(opts))
    eval(parse(text=paste0("options(",i,"=",opts[i],")")))
\end{lstlisting}
This pattern is a special case of code generation, recognizable by the
fact that \eval is executed in a loop.
